Connecting R with D3 for dynamic graphics, to explore multivariate data with tours

### 2018-03-07 paper-tourrd3.tex
small changes

The tour algorithm (Cook et al. 1995; Cook et al. 2007) is a way of
systematically generating and displaying projections of high-dimensional
spaces in order for the viewer to examine the multivariate distribution
of data. It can do this either randomly, or by picking projections
judged interesting according to some criterion or index function. The
tourr package (Wickham et al. 2011) provides the computing and display
in R (R Core Team 2018; Ihaka and Gentleman 1996) to make several types
of tours: grand, guided, little and local. The projection dimension can
be chosen between one and the number of variables in the data. The
display, though, has no capacity for interaction. The viewer can watch
the tour like a movie, but not pause it and restart, or change tour
type, or number of variables.

These interactive controls were provided with the tourrGui package
(Huang, Cook, and Wickham 2012), with was programmed with the RGtk2
package (Lawrence and {Temple Lang} 2010). This is not the toolkit of
choice today, and has been superceded with primarily web-capable tools,
like Shiny (Chang et al. 2017). To display dynamic graphics though, is
not straight-forward. This paper explains how to use D3 (Bostock,
Ogievetsky, and Heer 2011) as the display engine in a Shiny GUI, using
custom message passing between server and client.

@@

The main disadvantage is that the speed is inconsistent, as server and
client play tag to keep up with each other, and the display cannot
handle many observations. Noticeable slow down was oberved with 2000
points, the main reason being the rendering time required for the large
number of SVG circle elements. The situation can be improved when using
a single HTML5 canvas element to draw the scatter points, significantly
reducing the rendering time.

>>>>

The tour algorithm (Cook et al. 1995; Cook et al. 2007) is a way of
systematically generating and displaying projections of high-dimensional
spaces in order for the viewer to examine the multivariate distribution
of data. It can do this either randomly, or by picking projections
judged interesting according to some criterion or index function. The
tourr package (Wickham et al. 2011) provides the computing and display
in R (R Core Team 2018; Ihaka and Gentleman 1996) to make several types
of tours: grand, guided, little and local. The projection dimension can
be chosen between one and the number of variables in the data. The
display, however, has no capacity for interaction. The viewer can watch
the tour like a movie, but not pause it and restart, or change tour
type, or number of variables.

These interactive controls were provided with the tourrGui package
(Huang, Cook, and Wickham 2012), which was developed with the RGtk2
package (Lawrence and {Temple Lang} 2010). This is not the toolkit of
choice today, and has been superseded by primarily web-capable tools,
like Shiny (Chang et al. 2017). To display dynamic graphics though, is
not straightforward. This paper explains how to use D3 (Bostock,
Ogievetsky, and Heer 2011) as the display engine in a Shiny GUI, using
custom message passing between server and client.

@@

The main disadvantage is that the speed is inconsistent, as server and
client play tag to keep up with each other, and the display cannot
handle many observations. Noticeable slowdown was observed with 2000
points, mainly due to the rendering time required for the large
number of SVG circle elements. The situation can be improved when using
a single HTML5 canvas element to draw the scatter points, significantly
reducing the rendering time.
